On the Functionality of the Multilayer Perceptron

### 2020-12-22 Thesis/Drafts/Buell-PHYS799-ThesisDraft4.tex
Proofreading more Thesis

The task of recognizing the source of a sound wave in a digital audio file is trivial and almost effortless to a human being in many cases, but is far more difficult for a computer to do. In order to automate source source identification, we have designed and constructed a multimodal neural network that can classify audio files by what musical instrument the sound most resembles. The network uses two different input branches, one being a convolutional neural network (CNN), and the other, a multilayer perceptron (MLP). The CNN branch accepts input in the form of a two-dimensional spectrogram matrix, which encodes the energy distribution of an audio signal over frequency and time, and the MLP branch accepts input as a one-dimensional array of features derived from the physical nature of the sound wave. The spectrogram is produced by using a concatenation of multiple short-time Fourier transformations (STFT) of overlapping windowed time-space analysis frames. The feature vector implements techniques from digital signal processing to combine more explicit properties of the time-space and frequency-space information. To classify the audio files based on predicted sources, we aggregate and transform the activations from the final layers of each branch, and produce a single prediction label. We find that this architecture allows for reasonable classification performance, and demonstrates superior classification performance on average, when compared to either of it's constituent unimodal architectures. Training data for this experiment is from studio recordings of the Philharmonia Symphony Orchestra and the University of Iowa's Electronic Music Studios.

@@

\paragraph*{}The ability to classify time-series data into unique categories has become topic of much attention in the last few decades \cite{Khan,Mierswa}. From climate measurements, to financial reports, to audio-visual media, or to medicine, time-domain signals surround the modern world at every corner \cite{Virtanen}. In each context, the general idea is to place information into distinct categories called \textit{classes} based on inherent properties of the data. For example, doctors may wish to classify a patient as having a heart condition based on properties of their heartbeat or banks may wish to classify a purchase as fraudulent based on a history of spending habits. Often, the volume  of information or complexity of the task makes it unreasonable for a human to do without the aide of some form of automation.

\paragraph*{}It has long been a goal of computer scientists and engineers alike to develop a program, algorithm, or machine that could perform a certain task to the level of human-being \cite{Bishop,Geron,Loy}. While humans are intricate organisms capable of recognition or classification in many contexts, we do not have the ability to process millions of raw numerical values in the course of a few seconds. Similarly, modern computers are also complex systems that can filter through exabytes of information on a daily basis, but are not generally known for being able to distinguish cats from dogs \cite{Geron}. The compromise between humans and machines an algorithm called a \textit{neural network} that combines the computation speed of a computer with the decision-making power modeled from human brain \cite{Goodfellow,Levine,Loy}.

\paragraph*{}In this work, we have developed an automated neural network program that can classify digital sound waves as one of $37$ musical instruments that it most closely resembles. The general use of machine learning algorithms for audio information classification or identification is well explored, and a topic of much modern research \cite{Khan,Liu,Mierswa,Serizel,Zhang}. From previous work, it has been found that various types of audio classification are possible, and are still being expanded upon. Many papers detail possible sets of inputs for classifying speech against music or inputs for organizing audio-visual media based on content. In each case these inputs, called features, must often be tailored to meet the specific task at hand \cite{Virtanen,Liu}.

@@

\paragraph*{}As the name implies, a neural network is a computation a model or graph that is loosely structured off of the biological brain \cite{Geron,Loy}. Where a biological brain uses physical connects through axons to exchanges chemicals to communicate information, a neural network uses a chain of mathematical operations to communicate information \cite{Goodfellow}. The earliest work on attempting to model cognitive processes goes back to 1943, where Warren McCulloch and Walter Pitts began to develop a model of the human brain using principles of mathematics \cite{McCulloch,Loy}. Nearly ten years later, a simple neural networks called the \textit{perceptron} was developed by Frank Rosenblatt, which modeled connections of neurons through principles of linear algebra \cite{Geron,Petrik}. Another decade after that, inspiration from the visual cortex gave rise to the convolution operation to be applied to digital image processing.

@@

\paragraph*{}This project includes a number of steps from taking raw input data to the final classification program. In this section, we provide a very high level outline of the parts of this projects and show how each one influence the final product. Additionally, we indicate the sections in this thesis where the particular material is introduced and developed much more thoroughly.

@@

\paragraph*{}The smaller functions that comprise $F^*$ are called \textit{layers} \cite{Loy,Goodfellow}. Each type of layer performs a different operation and processes different types of input. In secs. (\ref{subsec-NerualNetworkIntro}) and (\ref{subsec-NetworkStructure}), we introduce the nature and structure of a neural network and how it can use used in part to solve this problem. We discuss it is inspired from a biological brain and show how different layer types are influenced from the brain's functions as well as how they are used in this problem. Using information about the problem, we carefully design the inputs, outputs and structure of the neural network function to be able to complete this task in a practical manner.

@@

\paragraph*{}In order to train the neural network to be able to identify musical instrument sources, we need a suitable data set to present to the model as training and validation data. The models uses this data and corresponding labels in the training process to modify it's function to allow for it to complete the task. University of Iowa Electronic Music Studio, and the London-Based Philharmonia Orchestra each have a large collection of publicly available audio files \cite{Philharmonia,UnivIowa}. These files contain short segments of musical instruments performing a single note, and are labeled according to the instrument that produced the sound. The files and labels make up the core of the data set that we use to train the model.

\paragraph*{}To ensure that these data sets are formatted consistently, we have designed a MATLAB program to read each audio file from it's original format, \textit{.AIF}, \textit{.MP3}, or similar, and rewrite the data as a new \textit{.WAV} files, sampled at 44.1 kHz with a 24-bit depth. This ensures that all data will have a consistent format when features are extracted. This stage is detailed in section (\ref{subsubsec-Preprocessing}). After the samples are formatted, we have produced a Python script with iterates through each new file and assigns a class label based on the title of the instrument. This has been design to account for the specific naming cases that arise. We assemble a dictionary-like structure of all training file with organizes the local file location of each sample and provides an integer and string label for all samples.

@@

\paragraph*{}The performance of a neural network is largely dependent of designing an appropriate set of predictors or features which represent the raw input given to the model. These inputs are designed to represent prolific properties of wave forms that can be expressed by a single or multiple numerical values. We detail these features in section (\ref{sec-Features}). We use tools from physics, mathematics, and signal processing to define and explore a comprehensive set of features that enables a high performance of the classifier.

@@

\paragraph*{}With the appropriate set of features designed, and the number of output categories determined, we can organize the structure of the neural network function. The shape of the network defines the flow of information and the hypothesis space. The nature of all possible solutions that the trained model could arrive at is subject to the architecture \cite{Goodfellow}. We construct a multimodal neural network, explored in section (\ref{subsec-Architecture}), that processes two distinct entry points derived from the same audio sample. This network is designed to evaluate each input independently, and concatenate the results to produce a single outputted prediction. This hybrid architecture combines the predictive power of the both a multilayer perceptron and a convolutional neural network. Our neural network was implemented using the tools available with the \textit{Tensorflow} module \cite{Tensorflow}.

@@

\paragraph*{} Because of the complexity of sound recognition, the challenge arises as to how to build some sort of program that could function at a level above general procedural or explicit instructional rules. Rather than \textit{hard-coding} a set of conditions or parameters for a classification program, we seek an solution that allows for a computer to \textit{learn} from labeled training examples in order \textit{teach} itself \cite{Bishop,Mitchell}. We choose to employ a neural network, which a computer algorithm that is designed to update itself to perform a task with increasing proficiency as it is presented with new information \cite{Geron2,Goodfellow,Levine,Mitchell}.

\paragraph*{}As the name implies, early neural networks were inspired from from the human brain. Former YouTube video classification team lead, and current machine learning consultant, Aurelien Geron writes about the relationship between biological brains and mathematical neural networks \cite{Geron}: 
\begin{quote}
Birds inspired us to fly, burdock plants inspired velcro, and nature has inspired many other inventions. It seems only logical, then, to look to the brain's architecture for inspiration on how to build an intelligent machine.
\end{quote}
The result is a computer program that is reminiscent of the a brain in that it structured much like the brain, and can\textit{learn} from new data. We model operations with collections are artificial neurons that are capable of connect and interacting with other neurons \cite{Levine}. Where a biological brain uses connections of chemical and electrical exchanges to process input and make decisions, the computational counterpart uses arithmetic and numerical exchanges to similarly process input and make decisions \cite{Levine,Bishop}. Where a brain uses physical connection of neurons through axons, the neural network uses mathematical connections of neurons through mathematical operations \cite{Geron,Levine}. We call the numerical value contained within an artificial neuron the \textit{activation} of that neuron \cite{Goodfellow,Geron}.

\paragraph*{}When a listener hears a sound, millions of neurons in their brain exchanges those chemical and electrical signals in such a way as to be able to identify the source of that sound \cite{Geron,Virtanen}. More conceptually, we can say that a listener is presented with a set of information $x$, which is air molecules vibrating near their tympanic membrane \cite{White,Olson}. That information is then passed into the brain where some function $F$ is used to model the incredibly complex set of neural connections. The output of that function is some label $y$, which the listener associates with a particular source \cite{Olson}. Similarly, we construct a neural network to accept a set of predictors $x$, which is passed into a function $F^*$ that is use to approximate $F$, and produce an output $y^*$, which predicts the source of the audio sample.

\paragraph*{}The neural network function $F^*$ uses a set of parameters $\Theta$, to map those inputs, $x$ to outputs, $y$ \cite{Goodfellow,James,Virtanen}. It is the process of updating the elements in the object $\Theta$ which allow us to train a neural network to perform the desired task. Similarly, it is the process of organizing and connecting neurons in the brain that allow humans to perform any specific task \cite{Geron,Levine}. Over the course of their lives, humans will be able to learn to match sounds to sources with relative ease, given the appropriate labels. In a similar manner, our neural network is presented a collection of predictors from sound files and a complementary set of labels indicating the source of each sound. This type of task is called \textit{classification}, which involves mapping the input data, $x$ to one of $k$ possible \textit{classes}, each represented by an integer, $0$ through $k-1$. Each of the $k$ classes represents a particular musical instrument or sound source that could have produced the sound-wave in the audio file \cite{Goodfellow,Loy,Virtanen}. In this section, we explore the motivation, structure, and training process of a neural network. We implement our neural network with the \textit{Tensorflow} Python package \cite{Tensorflow}.

@@

\paragraph*{}Step (ii.) is usually given by some \textit{activation function}:
\begin{equation}
\label{eqn-elementActivation}
x^{(l)} = \sigma^{(l)} \big[ z^{(l)} \big]
\end{equation}
Where $x^{(l)}$ is the final activations for layer $l$ and $z^{(l)}$ is given in equation (\ref{eqn-LinearTransform}). $\sigma^{(l)}$ is some activation function which introduces the ability to account for complicated non-linear decision boundaries which the neural network uses to make predictions. The combination of Eq. (\ref{eqn-LinearTransform}) and Eq. (\ref{eqn-elementActivation}), for a layer $l$ , make up the function represented by that layer, as in Eq. (\ref{eqn-altLayerFunction}). In this section, we detail the types of layer functions that are used to produce the classification model in this project.

@@

\paragraph*{}The Linear Dense Layer, often just called a \textit{Dense} Layer, or \textit{Fully-Connected} layer, was one of the earliest and most common function types used in artificial neural network models \cite{Geron2,Loy,McCulloch}. A dense layer is composed of a layer of \textit{artificial neurons}, each of which holds a numerical value within it, (usually a double-precision floating point number) called the \textit{activation} of that neuron. The activation of a dense layer are given by a linear combination of the layers inputs. This idea was developed from McCulloch and Pitts' work \cite{McCulloch} in attempting to model cognitive processes as mathematical functions by connecting neurons through a series of weighting values. The idea was expanded upon by Frank Rosenblatt in  1957 by combining many dense layers together to form a multilayer perceptron (MLP), which will make up a portion of the network that we have implemented. \cite{Geron,Loy,Levine}.

@@

\paragraph*{}Convolution layers emerged from studying the brain's visual cortex, and have been used from image recognition related tasks for around 40 years. \cite{Geron,Loy}. The layer function get's it's name from it's implementation of the mathematical \textit{convolution} operation. The 2D discrete-convolution of function or 2D array $A[i,j]$ and $B[i,j]$ is given by \cite{Goodfellow}:
\begin{equation}
\label{eqn-convolution}
C[i,j] = (A * B)[i,j] = \sum_{u}\sum_{v} A[i,j]B[i - u,j - v]
\end{equation}
Note that convolution is commutative: $(A * B) = (B * A)$. We implement a convolutional layer $f^{(l)}$ by creating a series of $K$ weighting matrices, called \textit{filters}, each of size $m \times n$, where $m,n \in \mathbb{Z}$. Often we choose $m = n$ and we call the shape the \textit{convolution kernel size} \cite{Loy, Goodfellow}. The step size in each of the two dimensions is known as the \textit{stride size}, which we choose to be $1$ in both dimensions.

@@

\paragraph*{}In general, for an $M \times N$ input, with kernel size $m \times n$, with stride size $1 \times 1$, and $K$ feature maps.
\begin{equation}
f^{(l)} : x^{(l-1)} \in \mathbb{R}^{(M \times N)} \rightarrow
x^{(l)} \in \mathbb{R}^{([M-m+1] \times [N-n+1] \times K)}
\end{equation}
As an example, consider a single filter map over input $x \in \mathbb{R}^{(3 \times 4)}$ and filter map $W \in \mathbb{R}^{(2 \times 2)}$ as shown in Fig. (\ref{fig-2DConvExample}):
\begin{figure}[H]
\begin{center}
	\begin{subfigure}{0.45\textwidth}
	\includegraphics[scale=0.6]{../figures/2DConvExample(a)}
	\caption{Input Activations, $x^{(l-1)}$}
	\end{subfigure}	
	\begin{subfigure}{0.45\textwidth}
	\includegraphics[scale=0.6]{../figures/2DConvExample(b)}
	\caption{Convolution Filter, $W^{(l)}_k$}
	\end{subfigure}	
	\begin{subfigure}{0.9\textwidth}
	\centering
	\includegraphics[scale=0.6]{../figures/2DConvExample(c)}
	\caption{Convolution Result, $x^{(l)}_k$}
	\end{subfigure}	
\end{center}
\caption{The result of convolving an input (a) with an filter map (b) is a new set of activations (c). This Image was adapted from Goodfellow, pg. 325 \cite{Goodfellow}}
\label{fig-2DConvExample}
\end{figure}
Note that formal implementations include a bias vector and an activation function.

@@

\paragraph*{}Pooling layers, such as maximum pooling, average pooling, or similar are typically placed after a layer or a set of layers of convolution \cite{Loy}. The purpose of this arrangement is to reduce the number of activations by selecting only the largest valus, thereby dropping the width of the model and ensuring only features with large activation values are preserved \cite{Geron,Loy,Goodfellow}. This layer "non-trainable" because it does not use any weights or transformations, only a fixed procedure that cannot be updated. Below, we detail what the "Call" method for a 2D maximum pooling layer might look like.

@@

\FOR {$i \in [0,1,...,M-2,M-1$}
	\FOR {$j \in [0,1,...,N-2,N-1$}
		\item Collect local activations in $P$, zero pad if necessary
		\item $P = x^{(l-1)}[i:i+m,j:j+n]$
		\item Find maximum value in $P$, and add to output
		\item $x^{(l)}[i,j] = \max\big( P \big)$
	\ENDFOR
\ENDFOR

@@

\paragraph*{}A flattening layer is used to compress an array with two or more dimensions down into a single dimension. For a flattening layer $(l)$, multidimensional activations in layer $(l-1)$ are rearranged down into an array such that each sample contains only one axis. Note that a batch of samples will still be two-dimensional. This is not like projecting the data into a lower dimension, but rather is the reorganization of values into an array of a row of column array. We can use function notation to express this for a single sample as:
\begin{equation}
\label{eqn-FlattenFunction}
f^{(l)} : x^{(l-1)} \in \mathbb{R}^{(M \times N \times ....)} \rightarrow
x^{(l)} \in \mathbb{R}^{(MN...\times 1)}
\end{equation}
The numerical value of each activation is left unchanged. For a layer with activation shape$M \times N$, the resulting activations are reshaped into $MN \times 1$ as shown in Eq.(\ref{eqn-FlattenFunction}).

\paragraph*{}Flattening Layer are most commonly used to prepare activations for entry into dense layer or series of dense layers. For example, an image may contain two or three dimension, and convolution over it will return two or three dimensional activations per sample. Each sample must be compressed into one dimensions so that it can be processed by dense layers. This layer also is "non-trainable" because it does no use any weights or transformations, only a fixed procedure.

@@

\paragraph*{}Activation functions are an important factor in the behavior of neural networks \cite{Geron}. For the convolution or dense layer, activation functions make up the second transformation step as in Eq. (\ref{eqn-elementActivation}) which allows for the neural network to produce non-linear decisions. In this section, we detail the activation functions used in this classification model.

@@

\paragraph*{}The Rectified Linear Unit (ReLU) activation function acts element-wise on the activations in a given layer. If the activation of a neurons is non-negative, the value is untouched, otherwise a $0$ is returned. For an input activation array $x$, ReLU is defined by:
\begin{equation}
\label{eqn-ReLU}
\text{ReLU}\big[x\big] = \max\big(0,x\big) = \left\{
\begin{array}{ll}
    x & \text{if\hspace{0.5cm}} x > 0 \\
    0 & \text{otherwise}
\end{array} 
\right.
\end{equation}
We provide a visualization of the function in Fig. (\ref{fig-ReLU}).
\begin{figure}[H]
\begin{center}
\includegraphics[scale=0.25]{../Figures/RectifiedLinearUnit}
\end{center}
\caption{Rectifed Linear Unit (ReLU) activation function}
\label{fig-ReLU}
\end{figure}
For our chosen architecture, ReLU is applied to the activations in every single Convolution layer and Dense Layer, with the exception of the output dense layer.

>>>>

The task of recognizing the source of a sound wave in a digital audio file is trivial and almost effortless to a human being in many cases, but is far more difficult for a computer to do. In order to automate audio source identification, we have designed and constructed a multimodal neural network that can classify audio files by what musical instrument the sound within it most resembles. The network uses two different input branches, one being a convolutional neural network (CNN), and the other, a multilayer perceptron (MLP). The CNN branch accepts input in the form of a two-dimensional spectrogram matrix, which encodes the energy distribution of an audio signal over frequency and time, and the MLP branch accepts input as a one-dimensional array of features derived from the physical nature of the sound wave. The spectrogram is produced by using a concatenation of multiple short-time Fourier transformations (STFT) of overlapping windowed time-space analysis frames. The feature vector implements techniques from digital signal processing to combine more explicit properties of the time-space and frequency-space information. To classify the audio files based on predicted sources, we aggregate and transform the activations from the final layers of each branch, and produce a single prediction label. We find that this architecture allows for reasonable classification performance, and demonstrates superior classification performance on average, when compared to either of it's constituent unimodal architectures. Training data for this experiment is from studio recordings of the Philharmonia Symphony Orchestra and the University of Iowa's Electronic Music Studios.

@@

\paragraph*{}The ability to classify time-series data into unique categories has become a topic of much attention in the last few decades \cite{Khan,Mierswa}. From climate measurements, to financial reports, to audio-visual media, or to medicine, time-domain signals surround the modern world at every corner \cite{Virtanen}. In each context, the general idea is to place information into distinct categories called \textit{classes} based on inherent properties of the data. For example, doctors may wish to classify a patient as having a heart condition based on properties of their heartbeat or banks may wish to classify a purchase as fraudulent based on a history of spending habits. Often, the volume  of information or complexity of the task makes it unreasonable for a human to do without the aide of some form of automation.

\paragraph*{}It has long been a goal of computer scientists and engineers alike to develop a program, algorithm, or machine that could perform certain tasks to the level of human-being \cite{Bishop,Geron,Loy}. While humans are intricate organisms capable of recognition or classification in many contexts, we do not have the ability to process millions of raw numerical values in the course of a few seconds. Similarly, modern computers are also complex systems that can filter through exabytes of information on a daily basis, but are not generally known for being able to distinguish cats from dogs \cite{Geron}. The compromise between humans and machines an algorithm called a \textit{neural network} that combines the computational speed of a computer with the decision-making power modeled from a biological brain \cite{Goodfellow,Levine,Loy}.

\paragraph*{}In this work, we have developed an automated neural network program that can classify digital sound waves as one of $37$ musical instruments that it most closely resembles. The general use of machine learning algorithms for audio information classification or identification is well explored, and is a steady topic of modern research \cite{Khan,Liu,Mierswa,Serizel,Zhang}. From previous work, it has been found that various types of audio classification are possible, and are still being expanded upon. Many papers detail possible sets of inputs for classifying speech against music or inputs for organizing audio-visual media based on content. In each case these inputs, called features, must often be tailored to meet the specific task at hand \cite{Virtanen,Liu}.

@@

\paragraph*{}As the name implies, a neural network is a computational a model or graph that is loosely structured off of the biological brain \cite{Geron,Loy}. Where a biological brain uses physical connects through axons to exchanges chemicals to communicate information, a neural network uses a chain of mathematical operations to communicate information \cite{Goodfellow}. The earliest work on attempting to model cognitive processes goes back to 1943, where Warren McCulloch and Walter Pitts began to develop a model of the human brain using principles of mathematics \cite{McCulloch,Loy}. Nearly ten years later, a simple neural networks called the \textit{perceptron} was developed by Frank Rosenblatt, which modeled connections of neurons through principles of linear algebra \cite{Geron,Petrik}. Another decade after that, inspiration from the visual cortex gave rise to the convolution operation to be applied to digital image processing.

@@

\paragraph*{}This project includes a number of steps from taking raw input data to the final classification program. In this section, we provide a very high level outline of the parts of this project and show how each one influences the final product. Additionally, we indicate the sections in this thesis where the particular material is introduced and developed much more thoroughly.

@@

\paragraph*{}We implement this function in the form of a \textit{neural network}. The smaller functions that comprise $F^*$ are called \textit{layers} \cite{Loy,Goodfellow}. Each type of layer performs a different operation and processes different types of input. In secs. (\ref{subsec-NerualNetworkIntro}) and (\ref{subsec-NetworkStructure}), we introduce the nature and structure of a neural network and how it can use used in part to solve this problem. We discuss its inspired from a biological brain and show how different layer types are influenced from the brain's functions as well as how they are used in this problem. Using information about the problem, we carefully design the inputs, outputs and structure of the function to be able to complete this task as needed.

@@

\paragraph*{}In order to train the neural network to be able to identify musical instrument sources, we need a suitable data set to present to the model as training and validation data. The model uses this data and corresponding labels in the training process to modify its function to allow for it to complete the task. University of Iowa Electronic Music Studio, and the London-Based Philharmonia Orchestra each have a large collection of publicly available audio files \cite{UnivIowa,Philharmonia}. These files contain short segments of musical instruments performing a single note, and are labeled according to the instrument that produced the sound. The files and labels make up the core of the data set that we use to train the model.

\paragraph*{}To ensure that these data sets are formatted consistently, we have designed a MATLAB program to read each audio file from it's original format, \textit{.AIF}, \textit{.MP3}, or similar, and rewrite the data as a new \textit{.WAV} files, sampled at 44.1 kHz with a 24-bit depth. This ensures that all data will have a consistent format when features are extracted. This stage is detailed in section (\ref{subsubsec-Preprocessing}). After the samples are formatted, we have produced a Python script which iterates through each new file and assigns a class label based on the name of the audio file. This has been design to account for the specific naming cases that may arise. We assemble a dictionary-like structure of all training files with organizes the local file location of each sample and provides an integer class label and string class label for each sample.

@@

\paragraph*{}The performance of a neural network is largely dependent on designing an appropriate set of predictors or features which represent the raw input given to the model. These inputs are designed to represent prolific properties of audio data that can be expressed by a single or multiple numerical values. We detail these features in section (\ref{sec-Features}). We use tools from physics, mathematics, and signal processing to define and explore a comprehensive set of features that enables a high performance of the classifier.

@@

\paragraph*{}With the appropriate set of features designed, and the number of output categories determined, we can organize the structure of the neural network function. The shape of the network defines the flow of information and the hypothesis space. The nature of all possible solutions that the trained model could arrive at is subject to the architecture \cite{Goodfellow}. We construct a multimodal neural network, explored in section (\ref{subsec-Architecture}), that processes two distinct entry points derived from the same audio sample. This network is designed to evaluate each input independently, and concatenate the results to produce a single outputted prediction. This hybrid architecture combines the predictive power of both a multilayer perceptron and a convolutional neural network. Our neural network was implemented using the tools available with the \textit{Tensorflow} module \cite{Tensorflow}.

@@

\paragraph*{} Because of the complexity of sound recognition, the challenge arises as to how to build some sort of program that could function at a level above general procedural or explicit instructional rules. Rather than \textit{hard-coding} a set of conditions or parameters for a classification program, we seek an solution that allows for a computer to \textit{learn} from labeled training examples in order \textit{teach} itself \cite{Bishop,Mitchell}. We choose to employ a neural network, which is a computer algorithm that is designed to update pieces of itself in order to perform a task with increasing proficiency as it is presented with new information \cite{Geron2,Goodfellow,Levine,Mitchell}.

\paragraph*{}As the name implies, early neural networks were inspired from from the human brain. Former YouTube video classification team lead, and current machine learning consultant, Aurelien Geron writes about the relationship between biological brains and mathematical neural networks \cite{Geron}: 
\begin{quote}
Birds inspired us to fly, burdock plants inspired velcro, and nature has inspired many other inventions. It seems only logical, then, to look to the brain's architecture for inspiration on how to build an intelligent machine.
\end{quote}
The result is a computer program that is structured much like the brain, and can\textit{learn} from new data. We model operations with collections of artificial neurons that are capable of connecting and interacting with other neurons \cite{Levine}. Where a biological brain uses connections of chemical and electrical exchanges to process input and make decisions, the computational counterpart uses arithmetic and numerical exchanges to similarly process input and make decisions \cite{Levine,Bishop}. Where a brain uses physical connection of neurons through axons, the neural network uses mathematical connections of neurons through operations \cite{Geron,Levine}. We call the numerical value contained within an artificial neuron the \textit{activation} of that neuron \cite{Goodfellow,Geron}.

\paragraph*{}When a listener hears a sound, millions of neurons in their brain exchanges those chemical and electrical signals in such a way as to be able to identify the source of that sound \cite{Geron,Virtanen}. More conceptually, we can say that a listener is presented with a set of information $x$, which is air molecules vibrating near their tympanic membrane \cite{White,Olson}. That information is passed into the brain where some function $F$ is used to model the incredibly complex set of neural connections. The output of that function is some label $y$, which the listener associates with a particular source \cite{Olson}. Similarly, we construct a neural network to accept a set of predictors $x$, which is passed into a function $F^*$ that is use to approximate $F$, and produce an output $y^*$, which predicts the source of the audio sample.

\paragraph*{}The neural network function $F^*$ uses a set of parameters $\Theta$, to map those inputs, $x$ to predicted outputs, $y^*$ \cite{Goodfellow,James,Virtanen}. It is the process of updating the elements in the object $\Theta$ which allow us to train a neural network to perform the desired task. Similarly, it is the process of organizing and connecting neurons in the brain that allow humans to perform a given task \cite{Geron,Levine}. Over the course of their lives, humans will be able to learn to match sounds to sources with relative ease, given the appropriate labels. In a similar manner, our neural network is presented a collection of predictors from sound files and a complementary set of labels indicating the source of each sound. This type of task is called \textit{supervised classification}, which involves mapping the input data, $x$ to one of $k$ possible \textit{classes}, each represented by an integer, $0$ through $k-1$. Each of the $k$ classes represents a particular musical instrument or sound source that could have produced the sound-wave in the audio file \cite{Goodfellow,Loy,Virtanen}. In this section, we explore the motivation, structure, and training process of a neural network. We implement our neural network with the \textit{Tensorflow} Python package \cite{Tensorflow}.

@@

\paragraph*{}Step (ii.) is usually given by some \textit{activation function}:
\begin{equation}
\label{eqn-elementActivation}
x^{(l)} = \sigma^{(l)} \big[ z^{(l)} \big]
\end{equation}
Where $x^{(l)}$ is the final activations for layer $(l)$ and $z^{(l)}$ is given in equation (\ref{eqn-LinearTransform}). $\sigma^{(l)}$ is some activation function which introduces the ability to account for complicated non-linear decision boundaries which the neural network uses to make predictions. The combination of Eq. (\ref{eqn-LinearTransform}) and Eq. (\ref{eqn-elementActivation}), for a layer $(l)$ , make up the function represented by that layer, as in Eq. (\ref{eqn-altLayerFunction}). In this section, we detail the types of layer functions that are used to produce the classification model in this project.

@@

\paragraph*{}The Linear Dense Layer, often just called a \textit{Dense} Layer, or \textit{Fully-Connected} layer, was one of the earliest and most common function types used in artificial neural network models \cite{Geron2,Loy,McCulloch}. A dense layer is composed of a collection of \textit{artificial neurons}, each of which holds a numerical value within it, (usually a double-precision floating point number) called the \textit{activation} of that neuron. The activations of a dense layer are given by a linear combination of the layers input activations. This idea was developed from McCulloch and Pitts' work \cite{McCulloch} in attempting to model cognitive processes as mathematical functions by connecting neurons through a series of weighting values. The idea was expanded upon by Frank Rosenblatt in  1957 by combining many dense layers together to form a multilayer perceptron (MLP), which will make up a portion of the network that we have implemented. \cite{Geron,Loy,Levine}.

@@

\paragraph*{}Convolution layers emerged from studying the brain's visual cortex, and have been used for image recognition related tasks for around 40 years. \cite{Geron,Loy}. The layer function get's it's name from it's implementation of the mathematical \textit{convolution} operation. The 2D discrete-convolution of function or 2D array $A[i,j]$ and $B[i,j]$ is given by \cite{Goodfellow}:
\begin{equation}
\label{eqn-convolution}
C[i,j] = (A * B)[i,j] = \sum_{u}\sum_{v} A[i,j]B[i - u,j - v]
\end{equation}
Note that convolution is commutative: $(A * B) = (B * A)$. We implement a convolutional layer $f^{(l)}$ by creating a series of $K$ weighting matrices, called \textit{filters}, each of size $m \times n$, where $m,n \in \mathbb{Z}$. We call the shape the \textit{convolution kernel size}, and often choose $m = n$ \cite{Loy, Goodfellow}. The step size in each of the two dimensions is known as the \textit{stride size}, which we choose to be $1$ in both dimensions.

@@

\paragraph*{}In general, for an $M \times N$ input, with kernel size $m \times n$, with stride size $1 \times 1$, and $K$ feature maps:
\begin{equation}
f^{(l)} : x^{(l-1)} \in \mathbb{R}^{(M \times N)} \rightarrow
x^{(l)} \in \mathbb{R}^{([M-m+1] \times [N-n+1] \times K)}
\end{equation}
As an example, consider a single filter map over input $x \in \mathbb{R}^{(3 \times 4)}$ and filter map $W \in \mathbb{R}^{(2 \times 2)}$ as shown in Fig. (\ref{fig-2DConvExample}):
\begin{figure}[H]
\begin{center}
	\begin{subfigure}{0.45\textwidth}
	\includegraphics[scale=0.6]{../figures/2DConvExample(a)}
	\caption{Input Activations, $x^{(l-1)}$}
	\end{subfigure}	
	\begin{subfigure}{0.45\textwidth}
	\includegraphics[scale=0.6]{../figures/2DConvExample(b)}
	\caption{Convolution Filter, $W^{(l)}_k$}
	\end{subfigure}	
	\begin{subfigure}{0.9\textwidth}
	\centering
	\includegraphics[scale=0.6]{../figures/2DConvExample(c)}
	\caption{Convolution Result, $x^{(l)}_k$}
	\end{subfigure}	
\end{center}
\caption{The result of convolving an input (a) with an filter map (b) is a new set of activations (c). This Image was adapted from Goodfellow, pg. 325 \cite{Goodfellow}}
\label{fig-2DConvExample}
\end{figure}
Note that formal implementations include a bias vector and an activation function.

@@

\paragraph*{}Pooling layers, such as maximum pooling, average pooling, or similar are typically placed after a layer or a set of layers of convolution \cite{Loy}. The purpose of this arrangement is to reduce the width of the model and ensure that only features with large activation values are preserved \cite{Geron,Loy,Goodfellow}. This layeris considered to be "non-trainable" because it does not use any weights or transformations, only a fixed procedure that cannot be updated. Below, we detail what the "Call" method for a 2D maximum pooling layer might look like.

@@

\FOR {$i \in [0,1,...,M-2,M-1]$}
	\FOR {$j \in [0,1,...,N-2,N-1]$}
		\item Collect local activations in $P$, zero pad if necessary
		\item $P = x^{(l-1)}[i:i+m,j:j+n]$
		\item Find maximum value in $P$, and add to output
		\item $x^{(l)}[i,j] = \max\big( P \big)$
	\ENDFOR
\ENDFOR

@@

\paragraph*{}A flattening layer is used to compress an array with two or more dimensions down into a single dimension. For a flattening layer $(l)$, multidimensional activations in layer $(l-1)$ are rearranged down into an array such that each sample contains only one axis. Note that a batch of samples will still be two-dimensional. This is not like projecting the data into a lower dimension, but rather is the reorganization of values into a row or column array. We can use function notation to express this for a single sample as:
\begin{equation}
\label{eqn-FlattenFunction}
f^{(l)} : x^{(l-1)} \in \mathbb{R}^{(M \times N \times ....)} \rightarrow
x^{(l)} \in \mathbb{R}^{(MN...\times 1)}
\end{equation}
The numerical value of each activation is left unchanged. For a layer with activation shape$M \times N$, the resulting activations are reshaped into $MN \times 1$ as shown in Eq.(\ref{eqn-FlattenFunction}).

\paragraph*{}Flattening Layer are most commonly used to prepare activations for entry into dense layer or series of dense layers. For example, an image may contain two or three dimensions, and convolution over it will return two or three dimensional activations per sample. Each sample must be compressed into one dimension so that it can be properly processed by dense layers. This layer also is "non-trainable" because it does not use any weights or transformations, only a fixed procedure.

@@

\paragraph*{}Activation functions are an important factor in the behavior of neural networks \cite{Geron}. For the convolution or dense layer, activation functions make up the second transformation step as in Eq. (\ref{eqn-elementActivation}) which allows for the neural network to model complicated non-linear decision boundaries that arise in real-world problems. In this section, we detail the activation functions used in this classification model.

@@

\paragraph*{}The Rectified Linear Unit (ReLU) activation function acts element-wise on the activations in a given layer. If the activation of a neuron is non-negative, the value is untouched, otherwise a $0$ is returned. For an input activation array $x$, ReLU is defined by:
\begin{equation}
\label{eqn-ReLU}
\text{ReLU}\big[x\big] = \max\big(0,x\big) = \left\{
\begin{array}{ll}
    x & \text{if\hspace{0.5cm}} x > 0 \\
    0 & \text{otherwise}
\end{array} 
\right.
\end{equation}
We provide a visualization of the function in Fig. (\ref{fig-ReLU}).
\begin{figure}[H]
\begin{center}
\includegraphics[scale=0.25]{../Figures/RectifiedLinearUnit}
\end{center}
\caption{Rectifed Linear Unit (ReLU) activation function}
\label{fig-ReLU}
\end{figure}
For our chosen architecture, ReLU is applied to the activations in every single Convolution layer and Dense Layer, with the exception of the output dense layer.
